\documentclass[man]{apa6}
\usepackage[american]{babel}
\usepackage[style=apa]{biblatex}
\usepackage{csquotes}
\DeclareLanguageMapping{american}{american-apa}
\bibliography{cites}
\title{Open Source Software and Innovation in the Software Development 
	Industry -- Draft}
\shorttitle{Open source software and innovation}
\author{Nathalie Rosalind Tate}
\affiliation{Oakland University}
\abstract{abstract text}

\begin{document}
\maketitle{}
		Open source software is software where the source code is freely available.
		Although the exact restrictions vary between open source licenses, most
		allow for users to edit and redistribute the software. One question raised
		by the rise of open source software (OSS) is how it compares to proprietary
		software with regard to innovation in the software industry, where
		``innovation'' is defined as a holistic metric that considers the amount of 
		new software and code commits, bug reports, and bug fixes to existing
		software, as well as the perceived quality and usability of said software.

		The main divide between proprietary software and OSS, other than the
		availability of the source code, is the price: Proprietary software
		frequently must be purchased, while OSS is usually provided for free. This
		price difference makes it easier to monetize proprietary software and is a
		large factor in the popularity of proprietary software.
		
		\section{Literature Review}
		
		\textcite{7} compares how price competition and
		differences in software quality vary between two markets: one in which two
		proprietary software firms are competing and one in which a proprietary
		software firm competes against an open source software (OSS) firm. The
		authors utilized a statistical model	to compare the two types of markets.
		The models compare software quality and price competition. Although OSS is
		often free, the authors factored in an ``effort cost'' when determining the
		price competition. The findings of the model were that a proprietary firm
		competing against an OSS firm may have either higher or lower profits than a
		proprietary firm competing against another proprietary firm, depending on
		the price competition from the OSS firm. The price competition depends on
		the effort cost of the OSS, the degree to which consumers value quality, and
		the degree to which the consumers of the OSS contribute to the production of
		the OSS. The proprietary firm had the highest profits when the price
		competition from the OSS firm was low, but the consumers were better-off
		with lower price competition. The authors also found that OSS had low value
		for users who valued quality in their software, but that it may have high
		value for users who did not value quality \parencite{7}.
		
		\textcite{10} discusses the visibility of software used
		and mentioned in biology literature. The authors compiled a ``balanced and
		representative sample'' of biology literature and analyzed the content by
		looking for mentions of	software. Once found, the mentions were ``coded''
		according to a coding scheme defined by the authors. The coding scheme
		contained information about	the mentioned software: if the mention was
		cited, if the mentioned software was easily identifiable and findable, if
		the source code was	available, and if the author of the software granted
		permission for others to edit it. The authors found that they had access to
		79\% of the mentioned	software, 47\% was available for free, 32\% allowed
		access to the source code, and 20\% allowed users to modify the code. The
		authors concluded that the manner in which software was mentioned and cited
		in scientific literature are ``varied and appear relatively ad hoc.'' The
		authors end with a call for more research \parencite{10}.

		\textcite{12} utilizes a statistical model to examine
		competition between a firm that produces proprietary software and a
		community that produces OSS. The author's findings are that the firm will
		cater to the less technologically savvy and the community will cater to
		those who are more technologically capable. In the presence of competition
		from the community, the firm targets ``even less technologically savvy
		individuals''. Furthermore, the author notes that the ``real price'' of
		software has two factors: the purchase price and the opportunity cost of
		learning how to use it. Additionally, the author comments on how proprietary
		firms can benefit from ``R\&D spillover'' from permissively licensed OSS
		projects \parencite{12}.

		\textcite{1} discusses the ``business of open source
		software''.  The authors discuss possible business models that allow an
		individual or corporation to profit from OSS. These business models include:
		``Support contracts and services", where one only supplies support for the
		software; ``Open core'', where the core software is OSS but the peripheral
		software and extensions are closed source; ``Business source'', where the
		software is maintained under two licenses with a time delay; ``Dual
		licensing'', where the software is maintained under a proprietary license
		and a strong copy-left license, so customers who cannot or chose not to
		comply with the terms of the copy-left license must buy the software; and
		``Software as a service'', where the application software is OSS, but the
		server it communicates with is closed source. \parencite{1}.

		\textcite{6} examines the effects of both free and
		commercial OSS on proprietary software. The author used a statistical model
		to compare two different hypothetical markets: one where proprietary
		software only competes with free OSS and another where proprietary software
		competes against both free and commercial OSS. The model considered
		differences in usability between OSS and proprietary software, as well
		differences in functionality and user technical ability. The authors
		findings were that the introduction of commercial OSS to the market can
		decrease the price of proprietary software and increase social welfare
		without necessarily decreasing the market share of the proprietary software.
		A market in which proprietary software only competes against free OSS,
		however, may lead to higher prices and lower social welfare \parencite{6}.

		\section{Methods}
		\section{Results}
		\section{Discussion}
		\textcite{7} compare a market where two proprietary firms compete to one in
		which a proprietary firm competes with a free OSS firm. 

		\printbibliography{}

\end{document}
